Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2023-02-16 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd
better cutoff histogram

# ROC analyses

@@

Due to low numbers of people with CFTR-related disorder or CRMS, this analysis focused on CF vs. non-CF until we receive additional data from NJH.

@@

### All participants

```{r}
# Exclude those with CFTR-related disorder or CRMS for now
roc_df = t1_df %>% filter(Diagnosis %in% c("Cystic Fibrosis","Non-CF"))
roc_df$Diagnosis = factor(roc_df$Diagnosis)
roc_df$Diagnosis = relevel(roc_df$Diagnosis,ref = "Non-CF")

roc_cf <- roc(response = roc_df$Diagnosis, predictor = roc_df$`Na/Cl Ratio`, quiet = T)
coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)

plot(roc_low, print.auc = T,print.thres=T)
```

The optimal Na/Cl threshold for distinguishing between CF and non-CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index, with and AUC of `r round(as.numeric(roc_cf$auc),3)`.

###

>>>>

# Cutoff analyses

@@

@@

### ROC

```{r}
# Re-format diagnosis column
t1_df$`Combined Diagnosis` <- factor(t1_df$Diagnosis,
                     levels = c("Non-CF","CFTR-related disorder","CRMS","Cystic Fibrosis"),
                     labels = c("Non-CF","CFTR-related disorder or CRMS","CFTR-related disorder or CRMS","Cystic Fibrosis"),
                     ordered = T)

```

```{r}
roc_cf <- roc(response= t1_df$`Combined Diagnosis`,
              predictor = t1_df$`Na/Cl Ratio`, 
              levels = c("CFTR-related disorder or CRMS","Cystic Fibrosis"),
              quiet = T)
coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)

roc_non_cf <- roc(response= t1_df$`Combined Diagnosis`,
              predictor = t1_df$`Na/Cl Ratio`, 
              levels = c("Non-CF","CFTR-related disorder or CRMS"),
              quiet = T)
coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)

plot(roc_cf, print.auc = TRUE, col = "blue")
plot(roc_non_cf, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
```

The blue curve shows the ROC for distinguishing between CFTR-related disorder or CRMS vs. CF, and the green is the ROC curve for distinguishing between non-CF vs. CFTR-related disorder or CRMS. The optimal threshold for distinguishing between CFTR-related disorder or CRMS vs. CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index. The optimal threshold for distinguishing between non-CF vs. CFTR-related disorder or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index. The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.

### Histogram by group

Vertical blue line indicates the threshold for CFTR-related disorder or CRMS vs. CF and green indicates the threshold for non-CF vs. CFTR-related disorder or CRMS. Na/Cl ratio was log-transformed for easier visualization.

```{r warning=FALSE}
ggplot(t1_df[!is.na(t1_df$`Combined Diagnosis`), ], 
       aes(log(`Na/Cl Ratio`), fill = `Combined Diagnosis`)) +
  geom_histogram(alpha = 0.3, position = "identity",bins = 50) +
  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
  geom_vline(xintercept = log(coords_non_cf$threshold), color = "green") +
  ylab("Count") + xlab("ln(Na/Cl Ratio)") +
  theme_bw()
```
